%!tex root=./thesis.tex
\vspace*{14cm}
\begin{center}
  \makeatletter
  \copyright\ \@author{} 2021\\All rights reserved
  \makeatother
\end{center}
\noindent
\begin{center}
  \footnotesize{~} %\aboutthesis
\end{center}
\noindent

\newpage

% \emph{
%     A thesis by compilation includes a signed declaration that specifies:

%     Title, authorship and publication outlet of each paper.
%     The current status of each paper (In press, Accepted, Under Review, In preparation).
%     The extent of the contribution of the candidate to the research and the authorship of each paper.

    % For each paper where the candidate is not the sole author, either:

    % The collaborating authors sign the declaration; or
    % A senior author signs the declaration on behalf of the collaborating authors

% }

\vspace*{4cm}
    I declare that the research presented in this thesis represents original
    work that I carried out during my candidature at the Australian National
    University, except for contributions to multi-author papers incorporated
    in the thesis where my contributions are specified in this Statement of
    Contribution.

    \begin{itemize}
      \item \emph{Radio Galaxy Zoo: Machine learning for radio source host galaxy cross-identification}, by \textbf{M. J. Alger}, J. K. Banfield, C. S. Ong, L. Rudnick, O. I. Wong, C. Wolf, H. Andernach, R. P. Norris, and S. S. Shabala. Published in 2018 in the \emph{Monthly Notices of the Royal Astronomical Society}. \autoref{cha:cross-id} in this thesis.
      \item \emph{Radio Galaxy Zoo: Radio luminosity functions of extended sources}, by \textbf{M. J. Alger}, O. I. Wong, C. S. Ong, N. M. McClure-Griffiths, H. Andernach, L. Rudnick, S. S. Shabala, A. F. Garon, J. K. Banfield, A. D. Kapi\'nska, R. P. Norris, and A. J. M. Thomson. Not yet submitted. \autoref{cha:rlfs} in this thesis.
      \item \emph{Interpretable Faraday Complexity Classification}, by \textbf{M. J. Alger}, J. D. Livingston, N. M. McClure-Griffiths, J. L. Nabaglo., O. I. Wong, and C. S. Ong. Accepted; to be published in \emph{Publications of the Astronomical Society of Australia}. \autoref{cha:faraday} in this thesis.
    \end{itemize}

    For all three papers I wrote the entirety of the content, produced all figures except where noted, and conducted all experiments. Other authors reviewed the papers, made suggestions, discussed ideas, and vitally contributed to the Radio Galaxy Zoo project.

\vspace*{4cm}

\hspace{8cm}\makeatletter\@author\makeatother\par
\hspace{8cm}\today
